---
title: "Calculate diffusion coefficients of sea water solutes"
author: "Marko Lipka"
date: "1/18/2018"
output: github_document
bibliography: library.bib
nocite: | 
  @Schulz2006, @Boudreau1997
---

The diffusivity $D_w$ of solutes in free water can be linearly interpolated from [temperature dependent theoretical values in de-ionized water](https://www.springer.com/de/book/9783540321439) and in [sea water](https://books.google.de/books/about/Diagenetic_models_and_their_implementati.html?id=kr0SAQAAIAAJ&redir_esc=y) for measured *in-situ* salinity and temperature.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message = F)

source("calculateDsal.R")
```

*calculate.Dsal(el, temperature, salinity)* returns the diffusion coefficient of dissolved species *el*  
 + Ca^2+^ ("Ca"),  
 + Mg^+^ ("Mg"),  
 + Na^+^ ("Na"),  
 + K^+^ ("K"),  
 + SO~4~^2-^ ("SO4"),  
 + Ba^2+^ ("Ba"),  
 + Fe^2+^ ("Fe"),  
 + Li^+^ ("Li"),  
 + Mn^2+^ ("Mn"),  
 + PO~4~^3-^ ("PO4"),  
 + HPO~4~^2-^ ("HPO4"),  
 + Sr^2+^ ("Sr"),  
 + H~2~S ("H2S"),  
 + HS^-^ ("HS"),  
 + HCO~3~^- ("HCO3"),  
 + H~4~SiO~4~ ("H4SiO4"),  
 + NH~4~ ("NH4"),  
 + NO~3~^-^ ("NO3"),  
 + CH~4~ ("CH4")  
 at a given *temperature* (in °C) and *salinity*.
 
```{r}
calculate.Dsal()
```
 
 