---
title: "Calculate diffusion coefficients of sea water solutes"
author: "Marko Lipka"
date: "1/18/2018"
output: github_document
bibliography: library.bib
---

The diffusivity *D* of solutes in sea water depends on temperature and salnity. It can be linearly interpolated from [temperature dependent theoretical values in de-ionized water](https://www.springer.com/de/book/9783540321439) and in [sea water](https://books.google.de/books/about/Diagenetic_models_and_their_implementati.html?id=kr0SAQAAIAAJ&redir_esc=y) for measured *in-situ* salinity and temperature.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message = F)

source("calculateDsal.R")
```

## Diffusivity calculation

*calculate.Dsal(el, temperature, salinity)* calculates for given *temperature* (in °C) and *salinity* the diffusion coefficient of a dissolved species *el*, which can be one of the following:  

|Solute|*el*|  |Solute|*el*|
|:-----|:---|-----------|:-----|:---|
| *Ca^2+^* | "Ca"|    | *PO~4~^3-^* | "PO4"|
| *Mg^+^* | "Mg"|     | *HPO~4~^2-^* | "HPO4"|
| *Na^+^* | "Na"|     | *Sr^2+^* | "Sr"|
| *K^+^* | "K"|       | *H~2~S* | "H2S"|
| *SO~4~^2-^* | "SO4"|| *HS^-^* | "HS"|
| *Ba^2+^* | "Ba"|    | *HCO~3~^-^* | "HCO3"|
| *Fe^2+^* | "Fe"|    | *NH~4~* | "NH4"|
| *Li^+^* | "Li"|     | *NO~3~^-^* | "NO3"|
| *Mn^2+^* | "Mn"|    | *CH~4~* | "CH4"|

 
```{r, include=FALSE}
elements <- c("Ba", "Ca", "CH4", "CO2", "Fe", "H2S", "HCO3", "HPO4", "HS", "K", "Li", "Mg", "Mn", "Na", "NH4", "NO3", "PO4", "Si", "SO4", "Sr")
```
 
 
 The function returns a list of parameters
  
```{r, echo = F}
res <- calculate.Dsal(el = "H2S", temperature = 12, salinity = 15)

names(res)
```
  
  + The input salinity (*sal*) and temperature (*t*) values.
  + Linear model coefficients *m0* and *m1* and calculated diffusivity *D~0~* at 0 salinity:  
  *D~0~(T) = (m0 + m1 \* T) \* 10^-6^ cm^2^ s^-1* [@Boudreau1997].
  + Interpolated diffusivity *D~35~* at a salinity of 35 from a temperature-diffusivity relation in the literature [@Schulz2006].
  + Calculated diffusivity *D~sal~* at the given salinity, linearly interpolated from *D~0~* and *D~35~*.

and three plots showing the calculation of the three diffusivities (*D~0~*, *D~35~* and *D~sal~*) via linear interpolation.
  
### Example
 
```{r example, echo = F, warning=F, fig.width=4, fig.height=3}
res$plot.D0
res$plot.D35
res$plot.Dsal
```
 
## Salinity and temperature dependency of molecular diffusivity

```{r, cache=TRUE, fig.width=10, fig.height=5}
for(el in elements){
    print(Dsal.TSdep(el))}
```
 
 
## Correction factors for stoichometric ratio calculations

For example diffusivity ratio of *H~2~S* / *SO~4~^2-^*:

```{r, cache=TRUE, fig.width=10, fig.height=5}
Dsal.element.ratios(ely = "H2S",
                    elx = "SO4")
```

# References